---
title: "Downloading data"
output: pdf_document
---
# Download relevant packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nhanesA)
library(knitr)
library(tidyverse)
library(SASxport)

```
#Decide year of dataset to use
#Find out what datasets I need to use- then use nhanes() to download them as dataframes
#Combine necessary datasets into one large one

```{r}
?nhanesTables
nhanesTables(data_group='DEMO', year = 2003)
nhanesTables(data_group='EXAM', year = 2003)
nhanesTables(data_group='LAB', year = 2003)
nhanesTables(data_group='Q', year = 2003)
nhanesTables(data_group= 'DIET', year = 2003)

nhanesTables(data_group='DEMO', year = 2005)
nhanesTables(data_group='EXAM', year = 2005)
nhanesTables(data_group='LAB', year = 2005)
nhanesTables(data_group='Q', year = 2005)
nhanesTables(data_group= 'DIET', year = 2005)

#get a description of the variables in the dataset
kable(nhanesTableVars(data_group = 'DEMO', nh_table = 'DEMO_C', namesonly = FALSE))
```
#Pull out each table that is relevant using search function on website
#The data file name is the table that you input here to retrieve it #and you also need the variable name
#to make sure you select just the variables you are interested in 
#Give it a name so that it can be merged later
```{r}
demo <- nhanes('DEMO_C')
bp <- nhanes('BPX_C')
bmi <- nhanes('BMX_C')
chol1 <- nhanes('DR1IFF_C')
chol2 <- nhanes('DR1TOT_C')
hdl <- nhanes('l13_c')
ldl <- nhanes('L13AM_C')
chol3 <- nhanes('L40_C')
smoke <- nhanes('SMQ_C')
pa <- nhanes('PAQ_C')
kidney <- nhanes('OHXREF_C')
diab <- nhanes('DIQ_C')
medical <- nhanes('MCQ_C')
mental <- nhanes('CIQGAD_C')
occupation <- nhanes('OCQ_C')
sleep <- nhanes('CIQDEP_C')

#tells you the name of the variables in the table
names(chol1)
```

```{r}
demo3 <- demo[c('SEQN', 'RIDRETH1', 'INDHHINC', 'RIAGENDR', 'RIDAGEYR', 'RIDAGEMN')]
demovars <- names(demo3)
demovars
#original dataset with changed column names
demo3 <- demo3 %>% rename(Number = SEQN, Ethnicity = RIDRETH1, Household_income = INDHHINC, Gender = RIAGENDR, Age = RIDAGEYR, Age_months = RIDAGEMN)
demo3
#translate data into categories and change column names
demo34 <- nhanesTranslate('DEMO_C', demovars, data = demo3)
demo34
```

```{r}
bmi3 <- bmi[c('SEQN', 'BMXBMI', 'BMXWT')]
bmivars <- names(bmi3)
bmivars
#original dataset with changed column names
bmi3 <- bmi3 %>% rename(Number = SEQN, BMI = BMXBMI, Weight = BMXWT)
bmi3
#translate data into categories and change column names
bmi34 <- nhanesTranslate('BMX_C', bmivars, data = bmi3)
bmi34
```
```{r}
bp3 <- bp[c('SEQN', 'BPXSY1', 'BPXDI1', 'BPXPULS')]
bpvars <- names(bp3)
bpvars
#original dataset with changed column names
bp3 <- bp3 %>% rename(Number = SEQN, Diastolic = BPXDI1, Systolic = BPXSY1, regularity = BPXPULS)
bp3
#translate data into categories and change column names
bp34 <- nhanesTranslate('BMX_C', bpvars, data = bp3)
bp34
```

```{r}
#chol13 <- chol1[c('SEQN', 'DR1ICHOL')]
#chol1vars <- names(chol13)
#chol1vars
#original dataset with changed column names
#chol13 <- chol13 %>% rename(Number = SEQN, Chol1 = DR1ICHOL)
#chol13
#translate data into categories and change column names
#chol134 <- nhanesTranslate('DR1IFF_C', chol1vars, data = chol13)
#chol134
```

```{r}

chol23 <- chol2[c('SEQN', 'DR1TCHOL')]
chol2vars <- names(chol23)
chol2vars
#original dataset with changed column names
chol23 <- chol23 %>% rename(Number = SEQN, Chol2 = DR1TCHOL)
chol23
#translate data into categories and change column names
chol234 <- nhanesTranslate('DR1TOT_C', chol2vars, data = chol23)
chol234
```

```{r}
names(hdl)
hdl3 <- hdl[c('SEQN', 'LBDHDDSI')]
hdl3vars <- names(hdl3)
hdl3vars
#original dataset with changed column names
hdl33 <- hdl3 %>% rename(Number = SEQN, hdl = LBDHDDSI)
hdl33
#translate data into categories and change column names
hdl334 <- nhanesTranslate('l13_c', hdl3vars, data = hdl33)
hdl334
```

```{r}
names(chol3)
chol33 <- chol3[c('SEQN', 'LBDSCHSI')]
chol3vars <- names(chol33)
chol3vars
#original dataset with changed column names
chol33 <- chol33 %>% rename(Number = SEQN, chol3 = LBDSCHSI)
chol33
#translate data into categories and change column names
chol334 <- nhanesTranslate('L40_C', chol3vars, data = chol33)
chol334
```

```{r}
names(diab)
diab3 <- diab[c('SEQN', 'DIQ010', 'DIQ050')]
diab3vars <- names(diab3)
diab3vars
#original dataset with changed column names
diab3 <- diab3 %>% rename(Number = SEQN, Diabetes = DIQ010, Insulin = DIQ050)
diab3
#translate data into categories and change column names
diab34 <- nhanesTranslate('L40_C', diab3vars, data = diab3)
diab34
```

```{r}
names(ldl)
ldl3 <- ldl[c('SEQN', 'LBDLDL')]
ldl3vars <- names(ldl3)
ldl3vars
#original dataset with changed column names
ldl3 <- ldl3 %>% rename(Number = SEQN, ldl = LBDLDL)
ldl3
#translate data into categories and change column names
ldl34 <- nhanesTranslate('L13AM_C', ldl3vars, data = ldl3)
ldl34
```

```{r}
names(smoke)
smoke3 <- smoke[c('SEQN', 'SMQ040', 'SMQ020', 'SMD650', 'SMD070', 'SMD057')]
smoke3vars <- names(smoke3)
smoke3vars
#original dataset with changed column names
smoke3 <- smoke3 %>% rename(Number = SEQN, smoking_status = SMQ040, smoked_100 = SMQ020, smoke_30days = SMD650, smoke_day = SMD070, number_quit = SMD057)
smoke3
#translate data into categories and change column names
smoke34 <- nhanesTranslate('SMQ_C', smoke3vars, data = smoke3)
smoke34
```

```{r}
names(pa)
pa3 <- pa[c('SEQN', 'PAD200', 'PAQ560', 'PAD460')]
pa3vars <- names(pa3)
pa3vars
#original dataset with changed column names
pa3 <- pa3 %>% rename(Number = SEQN, PA_previous = PAD200, number_activities = PAQ560, days_active = PAD460)
pa3
#translate data into categories and change column names
pa34 <- nhanesTranslate('PAQ_C', pa3vars, data = pa3)
pa34
```

```{r}
names(kidney)
kidney3 <- kidney[c('SEQN', 'OHQ144')]
kidney3vars <- names(kidney3)
kidney3vars
#original dataset with changed column names
kidney3 <- kidney3 %>% rename(Number = SEQN, kidney_disease = OHQ144)
kidney3
#translate data into categories and change column names
kidney34 <- nhanesTranslate('OHXREF_C', kidney3vars, data = kidney3)
kidney34
```

```{r}
names(medical)
medical3 <- medical[c('SEQN', 'MCQ160E', 'MCQ250G', 'MCQ160L', 'MCQ220', 'MCQ160F')]
medical3vars <- names(medical3)
medical3vars
#original dataset with changed column names
medical3 <- medical3 %>% rename(Number = SEQN, heart_attack = MCQ160E, relative_ha = MCQ250G, liver_problem = MCQ160L, cancer = MCQ220, stroke = MCQ160F)
medical3
#translate data into categories and change column names
medical34 <- nhanesTranslate('MCQ_C', medical3vars, data = medical3)
medical34
```

```{r}
#names(mental)
#mental3 <- mental[c('SEQN', 'CIQGPHE')]
#mental3vars <- names(mental3)
#mental3vars
##original dataset with changed column names
#mental3 <- mental3 %>% rename(Number = SEQN, mental_illness = CIQGPHE)
#mental3
##translate data into categories and change column names
#mental34 <- nhanesTranslate('CIQGAD_C', mental3vars, data = mental3)
#mental34
```

```{r}
names(occupation)
occupation3 <- occupation[c('SEQN', 'OCD240')]
occupation3vars <- names(occupation3)
occupation3vars
#original dataset with changed column names
occupation3 <- occupation3 %>% rename(Number = SEQN, occupation = OCD240)
occupation3
#translate data into categories and change column names
occupation34 <- nhanesTranslate('OCQ_C', occupation3vars, data = occupation3)
occupation34
```

```{r}
#names(sleep)
#sleep3 <- sleep[c('SEQN', 'CIQD025')]
#sleep3vars <- names(sleep3)
#sleep3vars
##original dataset with changed column names
#sleep3 <- sleep3 %>% rename(Number = SEQN, sleep_disturb = CIQD025)
#sleep3
##translate data into categories and change column names
#sleep34 <- nhanesTranslate('CIQDEP_C', sleep3vars, data = sleep3)
#sleep34
```

```{r}
#merge data tables together into a dataset
#use seqn (renamed as Number) to match files

nh2003 <- merge(bmi34, bp34, by = c('Number'), all = TRUE)
nh2003
#nh2003 <- merge(nh2003, chol134, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, chol234, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, chol334, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, demo34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, hdl334, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, medical34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, kidney34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, ldl34, by = c('Number'), all=TRUE)
#nh2003 <- merge(nh2003, mental34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, occupation34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, pa34, by = c('Number'), all=TRUE)
#nh2003 <- merge(nh2003, sleep34, by = c('Number'), all=TRUE)
nh2003 <- merge(nh2003, smoke34, by = c('Number'), all=TRUE)
nh2003

```

```{r}
#MERGE MORTALITY FILES

#install.packages("readr")    #if package is not installed already, then uncomment and run this line
#install.packages("dplyr")    #if package is not installed already, then uncomment and run this line
library(readr)
library(dplyr)


# the location where the .DAT file is saved:
setwd("C:/Users/charl/OneDrive/Desktop/Data Science/Semester 2/Dissertation/Code/Model Code")


# remove all objects from the R environment
#rm(list=ls())

################
#NHANES VERSION#
################

srvyin <- paste("NHANES_2003_2004_MORT_2015_PUBLIC.dat")   # full .DAT name here
srvyout <- "nhanes3mort" # shorthand dataset name here



# read in the fixed-width format ASCII file
dsn <- read_fwf(file=srvyin,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)

# create the ID (SEQN) for the NHANES surveys
dsn$seqn <- substr(dsn$publicid,1,5)
# NOTE:   SEQN is the unique ID for NHANES.

#Drop NHIS variables
dsn <- select(dsn, -publicid)
dsn <- select(dsn, -dodqtr)
dsn <- select(dsn, -dodyear)
dsn <- select(dsn, -wgt_new)
dsn <- select(dsn, -sa_wgt_new)


# Structure and contents of data
str(dsn)


# Variable frequencies

#ELIGSTAT: Eligibility Status for Mortality Follow-up
table(dsn$eligstat)
#1 = "Eligible"
#2 = "Under age 18, not available for public release"
#3 = "Ineligible"

#MORTSTAT: Final Mortality Status
table(dsn$mortstat, useNA="ifany")
# 0 = Assumed alive
# 1 = Assumed deceased
# <NA> = Ineligible or under age 18

#UCOD_LEADING: Underlying Cause of Death: Recode
table(dsn$ucod_leading, useNA="ifany")
# 1 = Diseases of heart (I00-I09, I11, I13, I20-I51)
# 2 = Malignant neoplasms (C00-C97)
# 3 = Chronic lower respiratory diseases (J40-J47)
# 4 = Accidents (unintentional injuries) (V01-X59, Y85-Y86)
# 5 = Cerebrovascular diseases (I60-I69)
# 6 = Alzheimer's disease (G30)
# 7 = Diabetes mellitus (E10-E14)
# 8 = Influenza and pneumonia (J09-J18)
# 9 = Nephritis, nephrotic syndrome and nephrosis (N00-N07, N17-N19, N25-N27)
# 10 = All other causes (residual)
# <NA> = Ineligible, under age 18, assumed alive, or no cause of death data

#DIABETES: Diabetes Flag from Multiple Cause of Death (MCOD)
table(dsn$diabetes, useNA="ifany")
# 0 = No - Condition not listed as a multiple cause of death
# 1 = Yes - Condition listed as a multiple cause of death
# <NA> = Assumed alive, under age 18, ineligible for mortality follow-up, or MCOD not available

#HYPERTEN: Hypertension Flag from Multiple Cause of Death (MCOD)
table(dsn$hyperten, useNA="ifany")
# 0 = No - Condition not listed as a multiple cause of death
# 1 = Yes - Condition listed as a multiple cause of death
# <NA> = Assumed alive, under age 18, ineligible for mortality follow-up, or MCOD not available

# Re-name the dataset, DSN, to the short survey name then remove other R objects
assign(paste0(srvyout), dsn)
rm(dsn, srvyin, srvyout)

# Combining datasets:
nhanes3mort <- nhanes3mort %>% rename(Number = seqn)
nhanes3mort
nh2003full <- merge(nh2003, nhanes3mort, by = c('Number'), all=TRUE)
```

```{r}
#Save dataset so it doesn't have to be run again
#Open by clicking File > Open > Select 'file.RData'
save(nh2003full, file = 'nh2003full.RData')

write.csv(nh2003full, file = 'nh2003full.csv')
glimpse(nh2003full)
```

```{r}
library(survival)

#removes all rows where there is 2 or 3 in ELIGSTAT column (not eligible for follow-up) or under 18 in age column
adult_nh <- subset(nh2003full, eligstat != 2 & eligstat != 3 & Age >= 18)
adult_nh

#survival object for when there is no missing mortality data- 5610 people
surv_adult <- Surv(time = adult_nh$permth_int, event = adult_nh$mortstat)
surv_adult


# or can just set up the formula in the model as Surv(time, event) and take the columns from there!
surv_adult

```

```{r}
#Save just adult file- open it using File > Open RData
save(adult_nh, file = 'adult_nh.RData')

write.csv(adult_nh, file = 'adult_nh.csv')
glimpse(adult_nh)
```

